---
title: "Data Journalism"
author: "Manh Ha Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

# Import library

```{r}
library(tidyverse)
library(mapview)
library(sf)
```

# List of all function

-   Note: all function are written based on the format of nsw_fuel structure.

## Plot gas station on map

-   Ensure columns correctly named long, lat, brand
-   Need further adjustment as station may sell many fuels type

```{r}
plot_points_on_map <- function(df, fueltype) {
  # Filter the data by fuel type
  df <- df[df$fueltype == fueltype, ]
  
  # Convert lat and long to spatial points
  points_sdf = st_as_sf(df, coords = c("long", "lat"), crs = 4326)
  
  # Plot the data
  mapview(points_sdf, zcol = "price", label = points_sdf$brand)
}

```

## Average fuel type price for given postcode

-   Ensure postcode, fueltype, price correctly named

```{r}
avg_fuel_price <- function(df, fueltype, postcode) {
  # Filter the data by fuel type
  df <- df[df$fueltype == fueltype, ]
  
  # Filter the data by postcode
  df <- df[df$postcode == postcode, ]
  
  # Calculate the average price
  avg_price <- mean(df$price)
  
  # Return the average price
  return(avg_price)
}

```

## Function to extract postcode from address

```{r}
# Extract postcode from address and mutate to postcode column
postcode <- function(df) {
  df <- df %>%
    mutate(postcode = str_extract(address, "\\d{4}"))
  return(df)
}
```

## Find closet gas station given a coordinates

-   Ensure long, lat, fueltype columns correctly named

```{r}
find_closest_gas_station <- function(df, fueltype, lat, long) {
  # Filter the data by fuel type
  df <- df[df$fueltype == fueltype, ]
  
  # Convert lat and long to spatial points
  points_sdf = st_as_sf(df, coords = c("long", "lat"), crs = 4326)
  
  # Convert lat and long to spatial points
  point_sdf = st_as_sf(data.frame(long = long, lat = lat), coords = c("long", "lat"), crs = 4326)
  
  # Find the closest gas station
  closest_gas_station <- st_nearest_feature(point_sdf, points_sdf)
  
  # Return the closest gas station (station code)
  return(closest_gas_station)
}

```

## Combine all csv files in a folder to a single dataframe (using parralel processing???)

```{r}
combine_csv <- function(path) {
  # Get all csv files in the folder
  files <- list.files(path = path, pattern = "*.csv", full.names = TRUE)
  
  # Combine all csv files into a single dataframe
  df <- do.call(rbind, lapply(files, read_csv))
  
  # Return the dataframe
  return(df)
}

```


```



