---
title: "Sample"
author: "Manh Ha Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

# Import lib

```{r}
library(tidyverse)
library(mapview)
library(leaflet)
library(sf)
```

# Import data

```{r}
nsw_fuel <- read.csv("nsw_fuelprices_2022-12-11.csv")
nt_fuel <- read.csv("nt_fuelprices_2022-12-11.csv")
qld_fuel <- read.csv("qld_fuelprices_2022-12-11.csv")
sa_fuel <- read.csv("sa_fuelprices_2022-12-11.csv")
tas_fuel <- read.csv("tas_fuelprices_2022-12-11.csv")
```

# All function

## Plot gas station on map

-   Ensure columns correctly named long, lat, brand
-   Need further adjustment as station may sell many fuels type

```{r}
plot_points_on_map <- function(df, fueltype) {
  # Filter the data by fuel type
  df <- df[df$fueltype == fueltype, ]
  
  # Convert lat and long to spatial points
  points_sdf = st_as_sf(df, coords = c("long", "lat"), crs = 4326)
  
  # Plot the data
  mapview(points_sdf, zcol = "price", label = points_sdf$brand)
}

```

## Average fuel type price for given postcode

-   Ensure postcode, fueltype, price correctly named

```{r}
avg_fuel_price <- function(df, fueltype, postcode) {
  # Filter the data by fuel type
  df <- df[df$fueltype == fueltype, ]
  
  # Filter the data by postcode
  df <- df[df$postcode == postcode, ]
  
  # Calculate the average price
  avg_price <- mean(df$price)
  
  # Return the average price
  return(avg_price)
}

avg_fuel_price(nsw_fuel, "U91", 2450)
```

## Find closet gas station given a coordinates

-   Ensure long, lat, fueltype columns correctly named

```{r}
find_closest_gas_station <- function(df, fueltype, lat, long) {
  # Filter the data by fuel type
  df <- df[df$fueltype == fueltype, ]
  
  # Convert lat and long to spatial points
  points_sdf = st_as_sf(df, coords = c("long", "lat"), crs = 4326)
  
  # Convert lat and long to spatial points
  point_sdf = st_as_sf(data.frame(long = long, lat = lat), coords = c("long", "lat"), crs = 4326)
  
  # Find the closest gas station
  closest_gas_station <- st_nearest_feature(point_sdf, points_sdf)
  
  # Return the closest gas station
  return(closest_gas_station)
}

station_code <- find_closest_gas_station(nsw_fuel, "U91", -30.28,153.12)
nsw_fuel[nsw_fuel$stationcode == station_code, ]
```

# Data wrangling

# Explore plotting

## NSW

Mutate postcode columns

```{r}
# Extract last 4 number of address as postcode

```

For all fuel

-   further away from SYD, price tend to be higher?

-   further in land, more expensive the fuel?

-   along the coast, relatively equal?

```{r}
fueltype <- unique(nsw_fuel$fueltype)
plot_points_on_map(nsw_fuel,fueltype[4]) 
```

Fuel type vs price

-   Reccommend using EV consistently cheaper than fossil fuel

```{r}
ggplot(nsw_fuel, aes(x=fueltype, y=price)) +
  geom_boxplot() +
  labs(title="Price Distribution by Fuel Type", x="Fuel", y="Price")

```

Price Distribution by Brand for a type of fuel

-   Observe that even for same type of fuel, and brand there can be variations between location

```{r}
# Combustion fuel price distribution  by brand
nsw_fuel %>%
  filter(fueltype != "EV") %>%   # remove EV as price is fixed
  filter(fueltype == "U91") %>%  # change fuel type
  ggplot(aes(x=brand, y=price)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Price Distribution by Brand", x="Brand", y="Price")

```

Fuel Type Popularity

-   Variations of petroleum are the most popular

```{r}
# Fuel Type Popularity
ggplot(nsw_fuel, aes(x=fueltype)) +
  geom_bar() +
  labs(title="Fuel Type Popularity", x="Fuel Type", y="Count")
```

How far do you have to travel on average to fill up your car???

-   Consider population density of each postcode to determine the weight of each postcode??

-   Randomly distributed a corresponding number of representative points in each postcode???

-   Calculate average distance from each point to the nearest gas station???

-   Consider the number of gas station in each postcode???

-   Problem: how to map postcode border??? + density within each postcode can fluctuate btw sections???



```{r}
# linear regression to fit price ~ long lat fueltype
nsw_model_parallel <- lm(price ~ long+lat+fueltype, data = nsw_fuel)
plot(nsw_model_parallel)
summary(nsw_model_parallel)

nsw_model_seperate <- lm(price ~ long*lat*fueltype, data = nsw_fuel)
nsw_model_seperate.step <- step(nsw_model_seperate, direction = "both")
summary(nsw_model_seperate.step)

BIC(nsw_model_parallel,nsw_model_seperate.step, nsw_model_seperate)
```
